---
title: "Short introduction"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Short introduction}
  %\VignetteEngine{knitr::knitr}
  %\VignetteDepends{TLMoments}
  %\VignettePackage{TLMoments}
  \usepackage[utf8]{inputenc}	
---

# TLMoments

`TLMoments` is a set of functions which main functionality is the calculation of Trimmed L-moments and their parameter and quantile estimates. One of the main goals is to reduce computation time compared to existing implementations (like `lmomco`, `Lmoments`, `Lmom`), therefore the core functions are written in C++ (see vignette "comparison of computation time" for speed comparisons). To ensure an easy usage, the package only contains a small set of functions. This vignette gives a short introduction to the most important ones and their usage. 

# Short Introduction

```{r echo=FALSE, message=FALSE}
library(TLMoments)
```

## Calculation of empirical TL-moments, parameter and quantile estimates. 

First we have a look at the basic functionality of calculating TL-moments and parameter and quantile estimates. Let assume we have a simple data vector:
```{r}
xvec <- evd::rgev(100, loc = 10, scale = 5, shape = .2)
```

To calculate TL-moments we can use the function `TLMoments` with arguments `leftrim`, `rightrim`, and `max.order` (generating an object of class `TLMoments`): 
```{r}
TLMoments(xvec)
TLMoments(xvec, leftrim = 0, rightrim = 1, max.order = 2)
```

We can generate parameters estimates by putting a `TLMoments`-object to the function `parameters` and specifying argument `distr`: 
```{r}
tlm <- TLMoments(xvec)
parameters(tlm, distr = "gev")

tlm <- TLMoments(xvec, rightrim = 1)
parameters(tlm, distr = "gev")
```

This generates an object of class `parameters`, which can be transmitted to `quantiles` to calculate quantile estimations: 
```{r}
tlm <- TLMoments(xvec)
quantiles(parameters(tlm, distr = "gev"), c(.9, .99, .999))

tlm <- TLMoments(xvec, rightrim = 1)
quantiles(parameters(tlm, distr = "gev"), c(.9, .99, .999))
```


## Support for different data types: 

These basic functions can not only be used for simple vectors of data, but for matrix-type data, list-type data, and data.frames as well. To demonstrate this, let's generate sample data of these four types: 
```{r}
xmat <- matrix(evd::rgev(100), nc = 4)
xvec <- xmat[, 3]
xlist <- lapply(1L:ncol(xmat), function(i) xmat[, i])
xdat <- data.frame(station = rep(1:4, each = 25), hq = as.vector(xmat))
```

The type of `lambdas` and `ratios` returned by `TLMoments` matches the input type: 
```{r}
TLMoments(xvec, leftrim = 0, rightrim = 1)
TLMoments(xmat, leftrim = 0, rightrim = 1)
TLMoments(xlist, leftrim = 0, rightrim = 1)
TLMoments(xdat, hq ~ station, leftrim = 0, rightrim = 1)
```

This holds when parameter and quantile estimations are calculated: 
```{r}
tlm <- TLMoments(xvec, leftrim = 0, rightrim = 1)
parameters(tlm, "gev")
tlm <- TLMoments(xmat, leftrim = 0, rightrim = 1)
parameters(tlm, "gev")
tlm <- TLMoments(xlist, leftrim = 0, rightrim = 1)
parameters(tlm, "gev")
tlm <- TLMoments(xdat, hq ~ station, leftrim = 0, rightrim = 1)
parameters(tlm, "gev")

tlm <- TLMoments(xvec, leftrim = 0, rightrim = 1)
quantiles(parameters(tlm, "gev"), c(.99, .999))
tlm <- TLMoments(xmat, leftrim = 0, rightrim = 1)
quantiles(parameters(tlm, "gev"), c(.99, .999))
tlm <- TLMoments(xlist, leftrim = 0, rightrim = 1)
quantiles(parameters(tlm, "gev"), c(.99, .999))
tlm <- TLMoments(xdat, hq ~ station, leftrim = 0, rightrim = 1)
quantiles(parameters(tlm, "gev"), c(.99, .999))
```

# Calculations using theoretical TL-moments and parameters

The functions `as.TLMoments` and `as.parameters` can be used to construct `TLMoments`- or `parameters`-objects of theoretical values (not calculated from data). These objects can be used in the same way like before (to convert between TL-moments and their parameters or to calculate the corresponding quantiles): 
```{r}
(tlm <- as.TLMoments(c(14.1, 4.3, 1.32)))
parameters(tlm, distr = "gev")
quantiles(parameters(tlm, distr = "gev"), c(.9, .99, .999))

(param <- as.parameters(loc = 10, scale = 5, shape = .2, distr = "gev"))
quantiles(param, c(.9, .99, .999))
TLMoments(param)
TLMoments(param, rightrim = 1)
```

Note, that we can simply use the `TLMoments`-function to calculate TL-moments corresponding to an `quantiles`-object. 

# Magrittr-Syntax

`TLMoments` is built to support the use in `magrittr`-Syntax. The nesting of functions can be written more readable as: 
```{r}
library(magrittr)

TLMoments(xvec, leftrim = 0, rightrim = 1) %>% 
  parameters("gev") %>% 
  quantiles(c(.99, .999))
```
